\textcolor{red}{Characterizing Water Quality with an Autonomous Robotic Team Employing Comprehensive In-Situ Sensing, Hyperspectral Imaging, Machine Learning, and Conformal Prediction}

### 2024-01-23 paper/paper.tex
updated paper text

%=================================================================
\documentclass[journal,article,submit,pdftex,moreauthors]{Definitions/mdpi}

@@

%=================================================================
% Full title of the paper (Capitalized)
\Title{\textcolor{red}{Leveraging Machine Learning to Characterize Water Quality with an Autonomous Team of Sensing Sentinels }}

% MDPI internal command: Title for citation in the left column
\TitleCitation{Robot Team II: Electric Boogaloo}

@@

% Authors, for the paper (add full first names)
\Author{John Waczak$^{\dagger}$\orcidA{}, Adam Aker, Lakitha O. H. Wijeratne, Shawhin Talebi, Bharana Fernando, Prabuddha Hathurusinghe, Mazhar Iqbal, David Schaefer, David J. Lary, $^{\dagger}$\orcidB{}*}

@@

% MDPI internal command: Authors, for citation in the left column
\AuthorCitation{Waczak, J.; Lary, D.; Lastname, F.}
% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
Hanson Center for Space Sciences, University of Texas at Dallas, Richardson, TX 75080, USA\
}

@@

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{\textcolor{red}{Advances in hyperspectral imaging technology are enabling significant improvements in the spectral resolution of remote sensing data products with a plethora of new satellites either in development or soon to be launched. Utilizing hyperspectral data for environmental assessment requires high quality in-situ data that is both difficult to collect and difficult to align with remote sensing data products due to orbital-dependent revisit times and occlusion by weather and cloud cover. In this study, present the results of a series of surveys.}}

% A robust characterization of model uncertainties is critical for any time-sensitive or dangerous application...To that end we employ a combination of model stacking together with Conformal Prediction by which we are able to blend together a variety of ML models while simultaneously estimating confidence intervals for the resulting predictions.

@@

% Keywords
\keyword{\textcolor{red}{Hyperspectral Imaging; Machine Learning; Robotic Teams; UAV;}}

@@

% Overview and Remote Sensing

% Water Quality

% UAVs

% Machine Learning

% Our Approach
- Most remote sensing data products are from multi-spectral imagers, i.e. LANDSAT, MODIS, ETC... 
    - multispectral imagers have limited spectral resolution 
    - Much progress has been made in the development of spectral indices
    - It can be challenging to directly associate these to desired measurements, rather we typically just track the index (without use of in-situ data), or generate classifications that can be challenging to interpret (i.e. good vs acceptable vs poor air/water quality)
- Existing hyperspectral imagers do not have global coverage, but many are on the way EnMAP
    - AVRIS was used in a plane 
    - Regardless, the increased data volume due to improved spectral resolution has an impact on the rate at which data are registered, processed, and transferred
- In situ measurements are expensive and challenging to collect
    - Existing in-situ data often require collection in the field and analysis back in the lab
    - In-situ measurements tend to have poor spatial resolution and volume (~10-100 of measurements per site)
    - Collection and analysis is slow. At best, we can have fixed location sensors which stream their data
- Retrieval of water quality parameters from remote sensing data requires the existence of co-aligned imagery
    - This could never be "real time" due to satellite revisit times (usual 1/day) 
    - remote sensing imagery are easily occluded by clouds and are impacted by atmospheric effects due to aerosols (dust) and gasses (ozone, water vapor, etc). 
- UAVs provided an excellent solution allowing for improved spatial resolution easily controlled by flight altitude
    - Much less expensive
    - Can rapidly survey wide areas that are hard to reach with conventional reference instruments
    - Can be equipped with many instruments including hyperspectral imagers
    - Can enable rapid development of relevant spectral indices by \textit{brining the spectrometer to the problem}
- UAVs alone are not a complete solution as we still suffer from an insufficient sampling density of in-situ reference data
    - The ideal solution is to use a team of autonomous sensing sentinels including a UAV to provide the hyperspectral imagery complemented by additional sentinels for collecting in-situ reference data
    - In this team, an uncrewed survey vehicle (USV), i.e. a boat is equipped with a suite of reference instruments. 
    - These instruments provide the ground truth data used to map reflectance measurements made by the UAV directly to the water quality parameters we actually care about.
- In our previous work, we devised and implemented such a team \cite{robotTeam1}. 
    - Prevously we showcased how the team can work for a single collection in the field 
    - In this work, we demonstrate how combining data from multiple collections allows the ML models to learn the representation of a wider range of target concentrations under a variety of lighting conditions. 
        - This is clearly beneficial when single-day collections are uni-modal. That is, for targets that show little variation within a single collection, there is a significant benefit to combining multiple days of observations 
        - This is successful even for collections with complicated distributions.
    - For use in near-real time safety assessments it's important that the proposed architecture be both fast and accurate
        - We describe the implementation of a rapid georectification procedure that can run \textit{on the UAV}
        - The selected ML models achieve a balance of both training time \textit{and} inference time allowing us to rapidly deploy the models
        - We additionally include a characterization of ML model uncertainty (which is typically lacking) using Conformal Prediction
        - Key reflectance features that impact model performance can be quickly identifed using permutation importance
            - This is less biased than typical tree-based impurity methods 
            - This is also model agnostic (can be applied to any black-box model) 
            - This is easily interpretable: we measure the impact on model performance ($R^2$ value) when a feature is randomly permuted and all others remain the same. The change in $R^2$ value tells us how sensivie the final model is to each feature 
            - This can aid the generation of new spectral indices as well as allow for reduced model size while maximizing performance, i.e. an Occam's razor approach. 
    - we showcase maps generated using the trained ML models

In our previous work, we developed a robotic team comprised of... We showed that using this team, hyperspectral imagery collected by the UAV can be calibrated against in-situ reference measurements collected by the USV to enable the rapid generation of wide area surveys for a variety of water quality parameters \cite{robotTeam1}. In this study, we expand on this previous work to demonstrate that by combining observations across multiple distinct collections, we can improve our machine learning models...

% The shrinking size of multispectral and hyperspectral imagers has enabled their inclusion as the payload of multi-rotor Unmanned Aerial Vehicles (quadcopters, hexacopters, etc.). These platforms are able to address the shortcoming of remote sensing imagery, namely poor spatial resolution and atmospheric interference effects due to aerosols, dust, water vapor, etc by flying close to the ground. Further, UAVs can be taken \textit{to the reference sensors} to better facilitate collocation with reference data. However,

% In-situ measurements are expensive to collect and suffer from poor spatial coverage. Remote sensing data cover much wider areas but do not directly provide parameters needed for environmental assessments. Additionally, remote sensing data, particularly satellite based imagers, also suffer from long orbit-dependent site revisit times and can be easily occluded by the presence of clouds. This makes it incredibly challenging to collocate sufficient quantities of in-situ measurements with overlapping remote sensing imagery to enable their calibration. \textcolor{red}{(add reference to David's CDOM paper here...)}. To address this gap, we have developed an autonomous team of sensing sentinels combining a mobile uncrewed surface vessel (USV) with a unmanned aerial vehicle (UAV) which together rapidly collect hyperspectral imgaery \textit{and} comprehensive in-situ measurements to enable the estimation of key environmental parameters including concentrations of key constituents like CDOM, Crude Oils, Ions,as well as physical parameters like temperature, salinity, and dissolved oxygen. In our previous work we established a methodology for calibrating the hyperspectral imagery against reference measurements during a collection window \cite{robotTeam1}. In his study, we demonstrate that such models can be reliably trained using data collected from multiple observations....

%The introduction should briefly place the study in a broad context and highlight why it is important. It should define the purpose of the work and its significance. The current state of the research field should be reviewed carefully and key publications cited. Please highlight controversial and diverging hypotheses when necessary. Finally, briefly mention the main aim of the work and highlight the principal conclusions. As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research. Citing a journal paper \cite{ref-journal}. Now citing a book reference \cite{ref-book1,ref-book2} or other reference types \cite{ref-unpublish,ref-communication,ref-proceeding}. Please use the command \citep{ref-thesis,ref-url} for the following MDPI journals, which use author--date citation: Administrative Sciences, Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, Journal of Intelligence, Journalism and Media, JRFM, Languages, Laws, Religions, Risks, Social Sciences, Literature.

@@

In this study we employ a team of autonomous sensing sentinels, hereafter referred to as \textit{the robot team}, to enable the rapid characterization of aquatic environments. This team consists of two key sensing sentinels: An uncrewed surface vessel (USV) used to collect in-situ reference measurements, and an unmanned aerial vehicle (UAV) for performing rapid, wide-area surveys to gather remote sensing data products. Both platforms are coordinated using the open-source QGroundControl software for flight control and mission planning and are equipped with high accuracy GPS and INS such that all collected data points are uniquely geolocated and time-stamped \cite{qgroundcontrol}. Both the USV and UAV are equipped with long-range Ubiquiti 5 GHz LiteBeam airMAX WiFi to enable streaming of data products to a ground station with network attached storage providing redundancy.

@@

The USV employed in the robot team is a Maritime Robotics Otter which features a modular design and 20 hour battery life. The USV is equipped with an in-situ sensing payload consisting of a combination of Eureka Manta-40 multi-probes including fluorometers, ion selective electrodes, and other physical sensors which together enable the collection of comprehensive near-surface measurements including Colored Dissolved Organic Matter (CDOM), Crude Oil, Blue-Green Algae (phycoerythrin and phycocyanin), Chlorophyll-A, $\mathrm{Na^+}$, $\mathrm{Ca^{2+}}$, $\mathrm{Cl^-}$, Temperature, Conductivity, and many others. The full list of measurements utilized in this study is given in Table~\ref{tab:sensors} and are categorized into four distinct types: physical measurements, ion measurements, biochemical measurements, and chemical measurements. Additionally, the USV is equipped with an ultra-sonic weather monitoring sensor for measuring air speed and direction as well as a a BioSonics MX Aquatic Habitat Echosounder sonar which are not explored in this study.

@@

\begin{table}[H] 
\caption{In-situ reference sensors modelled in this study.\label{tab:sensors}}
\newcolumntype{C}{>{\centering\arraybackslash}X}
\begin{tabularx}{\textwidth}{CCCC}
\toprule
\textbf{Sensor}	& \textbf{Units} & \textbf{Sensor Type} & \textbf{Target Category}\\
\midrule
Temperature		                    & $^{\circ}$C       & Thermistor  & Physical \\
Conductivity                        & $\mu$S/cm         & Four-Electrode Graphite Sensor & Physical  \\
%Dissolved Oxygen                    & mg/l              & Optical Sensor & Physical \\
pH                                  & logarithmic (0-14)& Flowing-junction Reference Electrode & Physical \\
Turbidity                           & FNU               & Ion-Selective Electrode  & Physical \\
$\mathrm{Ca^{2+}}$                  & mg/l              & Ion-Selective Electrode  & Ions \\
$\mathrm{Cl^-}$                     & mg/l              & Ion-Selective Electrode & Ions \\
$\mathrm{Na^+}$                     & mg/l              & Ion-Selective Electrode & Ions \\
Blue-Green Algae (phycoerythrin)    & ppb               & Fluorometer & Biochemical \\
Blue-Green Algae (phycocyanin)      & ppb               & Fluorometer & BioChemical \\
CDOM                                & ppb               & Fluorometer & Biochemical \\
Chlorophyll A                       & ppb               & Fluorometer & BioChemical \\
Optical Brighteners                 & ppb               & Fluorometer & Chemical \\
Crude Oil                           & ppb               & Fluorometer & Chemical\\
%Salinity                            & PSU               & Four-Electrode Graphite Sensor \\
%$\mathrm{NO_3^-}$                   & mg/l as nitrogen  & Ion-Selective Electrode \\
\bottomrule
\end{tabularx}
% \noindent{\footnotesize{\textsuperscript{1} Tables may have a footer.}}
\end{table}

As shown in Table~\ref{tab:sensors}, the physical parameter and ion sensors are largely based on different configurations of electrodes while the chemical and biochemical parameters are all optically active substances enabling their determination by fluorometry. The pigments phycoerythrin and phycocyanin are used to determine blue-green algae content which together with chlorophyll-a enables us to asses the distribution of photosynthetic life. Additionally, we also measure the concentration of colored dissolved organic matter (CDOM) which impacts light penetration and serves as a primary source of bioavailable carbon. Crude oil (natural petroleum) and optical brightener concentrations are also measured with fluorometers and are relevant for identifying sources of industrial contamination or natural seepage.

@@

\begin{figure}[H]
\begin{adjustwidth}{-\extralength}{0cm}
\centering
\includegraphics[width=16cm]{paper/figures/materials-and-methods/pipeline-figure-2.pdf}
\end{adjustwidth}
\caption{Hyperspectral Image Processing: Hyperspectral datacubes are collected one scan-line at a time (left). By utilizing downwelling irradiance spectra we convert each pixel from spectral radiance to reflectance. By using orientation and position data from the on-board GPS and INS, we georeference each pixel to assign it a latitude and longitude on the ground. The final data product is a georectified hyperspectral reflectance data cube visualized at the right. \label{fig:hsi-pipeline}}
\end{figure}

The hyperspectral imager utilized in our robot team is in a so-called \textit{pushbroom} configuration, that is, each image captured by the drone is formed one scan-line at a time as the UAV flies. Each scan-line consists of 1600 pixels for which incoming light is diffracted into 462 wavelength bins. In the collection software, a regular cutoff of 1000 lines is chosen so that each resulting \textit{image} forms a array of size 462$\times$1600$\times$1000 called a \textit{hyperspectral data cube}. Initially, the sampled spectra are in units of spectral radiance (measured in microflicks) however this does not account for the variability of incident light illuminating the water. To this end, we convert the hyperspectral data cubes into units of reflectance by utilizing the skyward-facing downwelling irradiance spectrometer (see Figure~\ref{fig:drone-components}). When our camera is normal to the water surface, then the reflectance is given by
\begin{equation}
    R(\lambda) = \pi L(\lambda)/E_d(\lambda)
\end{equation}
where $L$ is the spectral radiance, $E_d$ is the downwelling irradiance, and a factor of $\pi$ steradians results from assuming the water surface is Lambertian (diffuse) \cite{reflectance-conversion}. We will address the impact of this assumption in Section~\ref{sec:discussion}.

Having converted the hyperspectral data cube to units of reflectance, we must also georeference each pixel into a geographic coordinate system so that each image pixel can be assigned a latitude and longitude corresponding to the location on the ground from which it was sampled. During our three surveys, the UAV was flown at an altitude of ~50 m above the water. At this scale, the surface is essentially flat so that the hyperspectral data cube can be reliably georectified without the need for an on-board digital elevation map (DEM). We adopt the approach outlined in \cite{GeorectificationMuller, GeorectificationBaumker, GeorectificationMostafa} whereby each scan-line is georeferenced using the known field of view (30.8 $\deg$) together with the position and orientation of the UAV as provided by the on-board GPS INS. After a sequence of coordinate transformations, pixel coordinates are obtained in the relevant UTM zone (in meters). The resulting image is then re-sampled to a final output resolution. For these collections a resolution of 10 cm was utilized, however this can be adjusted to optimize the processing time and final resolution for real-time applications. Finally, the obtained pixel UTM coordinates are transformed to latitude and longitude for easy comparison with in-situ data collected by the USV. The final result is a georectified hyperspectral reflectance data cube. In Figure~\ref{fig:hsi-infographic} we visualize one such data cube, highlighting a selection of exemplar pixel spectra from the scene as well as the incident downwelling irradiance spectrum. A pseudo-color image is generated (plotted on the top of the data cube) to illustrate the scene.

@@

The above processing pipeline was implemented using the Julia programming language, a just-in-time compiled language with native multi-theading support \cite{julia-1}. By running this pipeline on the on-board processing computer, we are able to process the collected hyperspectral data cubes in near real time. This feature is critical for time-sensitive applications where we need to quickly asses if an area is safe and can not afford to wait to download and post-process collected imagery \textit{after} a flight.

\subsection{Data Collection and Training Set Curation}

Data were collected using the robot team at a site in north Texas across three observation periods in the fall of 2020. For each collection period, the UAV first completed two wide surveys of the lake capturing multiple hyperspectral data cubes. Subsequently, the the USV sampled across the lake, collecting in-situ reference measurements. Each of these reference measurements was then collocated with individual pixel spectra where the USV track overlapped with UAV pixels. To account for any time lag in the values measured by the in-situ instruments and to account for the USV size in comparison to the data cube spatial resolution, each in-situ measurement is associated with a 3$\times$3 grid of HSI pixels, that is, a 30 cm $\times$ 30 cm square. These combined data form the tabular data set on which we train regression models; pixel spectra form input features and each separate USV sensor forms a distinct target variable.

Each collection was performed near solar noon to maximize the amount of incident sunlight illuminating the water, however, to account for any potential variation in lighting conditions between data cubes, we augment the training set with additional features including the drone viewing geometry (roll, pitch, heading) and solar illumination geometry (solar azimuth, solar elevation, solar zenith) as well as the total at-pixel intensity (before refelctance conversion), the total downwelling intensity, and the drone altitude. Further feature engineering is performed to add a additional spectral indices which utilize combinations of specific wavelength bands like the normalized difference vegetation index (NDVI), normalized difference water index (NDWI), simple ratio (SR), photochemical reflectance index (PRI), and more as described in \cite{envi_vegetation_indices, thenkabail2018hyperspectral,kaufman1992atmospherically, SpectralIndexWheat}. A comprehensive list of these added features is provided in supplementary table S1. The final dataset includes a total of 526 features (462 reflectance bands plus 64 additional features) with over 120,000 records.

@@

Using these tools, the training procedure for each model is as follows: First, the model is trained using 6-fold cross validation on the training set with default hyperparameters. Feature importances and for the trained model are then computed and the top 25 features are identified. A second, reduced-feature model is then trained using the same 6-fold cross validation scheme. Next, each model is hyperparameter optimized using a random parameter search over the number of tees and sampling fraction. Model performance is evaluated by computing out-of-fold values for the coefficient of determination, as well as
\begin{align}
    \text{RMSE} &= \sqrt{\frac{\sum\limits_{i=1}^N (\hat{y}_i-y_i)^2}{N}}, \\
    \text{MAE} &= \frac{\sum\limits_{i=1}^N \lvert \hat{y}_i - y_i \vert}{N},
\end{align}
where RMSE is the root-mean-square error, and MAE is the mean absolute error between true values $y_i$ and predictions $\hat{y}_i$.

>>>>

%=================================================================
\documentclass[sensors,article,submit,pdftex,moreauthors]{Definitions/mdpi}

@@

%=================================================================
% Full title of the paper (Capitalized)
 \Title{\textcolor{red}{Characterizing Water Quality with an Autonomous Robotic Team Employing Comprehensive In-Situ Sensing, Hyperspectral Imaging, Machine Learning, and Conformal Prediction}}
%\Title{\textcolor{red}{Towards Real-Time Water Quality Characterization Using an Autonomous Robotic Team Employing Comprehensive In-Situ Sensing, Hyperspectral Imaging, and Conformal Prediction}}

% MDPI internal command: Title for citation in the left column
%\TitleCitation{Robot Team II: Electric Boogaloo}
\TitleCitation{\textcolor{red}{Characterizing Water Quality with an Autonomous Robotic Team Employing Comprehensive In-Situ Sensing, Hyperspectral Imaging, Machine Learning, and Conformal Prediction}}

@@

% Authors, for the paper (add full first names)
\Author{John Waczak\orcidA{}, Adam Aker, Lakitha O. H. Wijeratne, Shawhin Talebi, Bharana Fernando, Prabuddha Hathurusinghe, Mazhar Iqbal, Matthew Lary, David Schaefer, David J. Lary\orcidB{}*}

@@

% MDPI internal command: Authors, for citation in the left column
\AuthorCitation{\textcolor{red}{Waczak, J. et al.}}
% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address[1]{%
Hanson Center for Space Sciences, University of Texas at Dallas, Richardson, TX 75080, USA;  John.Waczak@utdallas.edu (J.W.); Adam.Aker@utdallas.edu (A.A.); lhw150030@utdallas.edu (L.O.H.W.); Shawhin.Talebi@utdallas.edu  (S.T.); ashen.fernando@utdallas.edu  (B.F.); prabuddhamadusanka.hathurusinghedewage@utdallas.edu (P.H.); mazhar.iqbal@utdallas.edu (M.I.); mdlary@me.com~(M.D.L.); captdaveschaefer@gmail.com (D.S.); david.lary@utdallas.edu (D.J.L.)}
\corres{\hangafter=1 \hangindent=1.05em \hspace{-0.82em} Correspondence: {David.Lary@utdallas.edu}}

@@

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{\textcolor{red}{
Inland waters pose a unique challenge for water quality measurement by remote sensing techniques due to their complicated spectral features and small-scale variability. At the same time, collecting the high-quality reference data needed to calibrate remote sensing data products is both time-consuming and expensive. In this study, we present further development of a robotic team composed of an uncrewed surface vessel (USV) providing in-situ reference measurements and an unmanned aerial vehicle (UAV) equipped with a hyperspectral imager. Together, this team is able to address the limitations of existing approaches by enabling the simultaneous collection of comprehensive in-situ data together with high resolution hyperspectral imagery. We showcase the capabilities of this team using data collected in a North Texas pond during three observations in the Fall of 2020. Machine learning models for 13 parameters are trained using the dataset of paired in-situ measurements and coincident reflectance spectra. Using conformal prediction we  estimate 90\% confidence intervals for the output of each trained model. Maps generated by applying the models to the collected imagery reveal the small-scale spatial variability within the pond. Additionally, computing the permutation importance for each reflectance bin provides further insight into the relevant spectral features associated with each measurement.
}}

@@

% Keywords
\keyword{Water Quality; Robotic Teams; Hyperspectral Imaging; Machine Learning; Conformal Prediction}

@@

For decades, remote sensing imagery have been used for environmental monitoring with applications ranging from resource mapping, land type classification, and urban growth assessment to wildfire tracking, natural disaster tracking, and many more \cite{cohen2004landsat, melesse2007remote, allison2016airborne, joyce2009review, robert2017analysis}. Among these applications, the retrieval of water quality parameters from remote sensing imagery remains challenging due to the difficulty of obtaining in-situ reference data coincident with available satellite imagery. Many studies have shown success modeling relevant water quality parameters such as colored dissolved organic matter (CDOM), cholorophyll-a, and total suspended sediment concentrations by using combinations of spectral bands from terrestrial remote sensing satellites. However, the in-situ reference data used in these studies typically require the time-consuming collection of individual samples for lab analysis or continuous monitoring at fixed sites \cite{brezonik2005landsat, paavel2011optical, gursoy2019investigating, bonansea2019using, sagan2020monitoring, remote-sensing-finland}.

Much effort has recently gone into the curation of comprehensive datasets combining water quality records with decades of satellite imagery to enable the development of new methods for water quality estimation. For example, Aurin et al. curated over 30 years of oceanographic field campaign data with associated coincident satellite imagery \cite{aurin2018remote}. Similarly, Ross et al. have combined over 600,000 records of dissolved organic carbon, chlorophyll-a, and other water quality parameters together with historical reflectance data from Landsat for the period of 1984-2019 \cite{ross2019aquasat}. With a sufficient quantity of data, machine learning methods can augment traditional approaches using reflectance band ratios to enable estimation of water quality metrics and perform water quality classification \cite{schiller1999neural, brown2008neural, lary2010artificial, feng2015urban,belgiu2016random}. However, the low spatial and spectral resolution of available multi-band remote sensing satellites makes it difficult to analyze inland waters with small features and complicated spectral signatures. Additionally, the long revisit times of these satellites limits their use for real-time monitoring, for example, during oil spills \cite{fingas2017review}.

Advances in multispectral and hyperspectral imaging technology have considerably reduced their size making it possible to incorporate these cameras into the payload of autonomous aerial vehicles (UAV) such as quadcopters, hexacopters, and other drones \cite{hruska2012radiometric}. By controlling the altitude during flight, UAVs can produce imagery with pixels at the centimeter scale and limit the need for complicated atmospheric corrections to account for scattering by atmospheric aerosols and gases \cite{adao2017hyperspectral,banerjee2020uav}. Already, UAVs equipped with multi- and hyperspectral imagers are being utilized in a variety of domains to great effect including biomass estimation, forest management, precision agriculture, and now, water quality monitoring \cite{adao2017hyperspectral, padua2017uas,arroyo2019implementation,kurihara2020unmanned,ehmann2019monitoring, lu2021retrieval}.

Despite the increased spectral and spatial resolution afforded by UAV platforms, these improvements alone are unable to address the limited spatial resolution of in-situ reference data.  To address this gap and enable real-time comprehensive water quality sensing, we have developed a robotic team composed of an autonomous uncrewed surface vessel (USV) equipped with a collection of reference grade instruments together with an UAV carrying a hyperspectral imager \cite{robotTeam1}. By incorporating these reference instruments on a maneuverable USV platform, we are able to rapidly collect large volumes of water quality data for a variety of physical, biochemical, and chemical parameters. Critically, the USV enables the collection of reference data with significantly improved spatial resolution compared to other approaches. These in-situ measurements then provide the ground-truth data used to train machine learning models which map reflectance spectra captured by the hyperspectral imager to the desired water quality parameters.

In this study, we expand on our previous work by demonstrating the capabilities of this coordinated robot team using data collected at a pond in North Texas across three separate days in the Fall of 2020. Specifically, we showcase the ability to accurately predict physical parameters such as temperature, conductivity, pH, and turbidity, as well as biochemical constituents including blue-green algae pigments, CDOM, and chlorophyll-a, in addition to concentrations of crude oils, optical brighteners, and a variety of ions. Further, the considerable size of the collected dataset allows us to quantify the uncertainty in the resulting machine learning models by using techniques from conformal prediction. This methodology addresses the shortcomings of existing approaches by facilitating rapid, real-time data collection coupled with on board hyperspectral image processing and machine learning model deployment with uncertainty quantification.

% critical to this approach

@@

 The robotic team employed in this study consists of two key sensing sentinels: An uncrewed surface vessel (USV) used to collect in-situ reference measurements, and an unmanned aerial vehicle (UAV) for performing rapid, wide-area surveys to gather remote sensing data products. Both platforms are coordinated using the open-source QGroundControl software for flight control and mission planning and are equipped with high accuracy GPS and INS such that all collected data points are uniquely geo-located and time-stamped \cite{qgroundcontrol}. Both the USV and UAV include long-range Ubiquiti 5 GHz LiteBeam airMAX WiFi to enable streaming of data products to a ground station with network attached storage providing redundancy.

@@

The USV employed in the robot team is a Maritime Robotics Otter which features a modular design and 20 hour battery life. The USV is equipped with an in-situ sensing payload consisting of a combination of Eureka Manta-40 multi-probes including fluorometers, ion selective electrodes, and other physical sensors which together enable the collection of comprehensive near-surface measurements including Colored Dissolved Organic Matter (CDOM), Crude Oil, Blue-Green Algae (phycoerythrin and phycocyanin), Chlorophyll-A, $\mathrm{Na^+}$, $\mathrm{Ca^{2+}}$, $\mathrm{Cl^-}$, Temperature, Conductivity, and many others. The full list of measurements utilized in this study is given in Table~\ref{tab:sensors} and is categorized into four distinct types: physical measurements, ion measurements, biochemical measurements, and chemical measurements. Additionally, the USV is equipped with an ultra-sonic weather monitoring sensor for measuring air speed and direction as well as a a BioSonics MX Aquatic Habitat Echosounder sonar which are not explored in this study.

@@

\begin{table}[H] 
\caption{In-situ reference sensors modelled in this study.\label{tab:sensors}}
\newcolumntype{C}{>{\centering\arraybackslash}X}
\begin{tabularx}{\textwidth}{CCCCC}
\toprule
\textbf{Sensor}	& \textbf{Units} & Resolution & \textbf{Sensor Type} & \textbf{Target Category}\\
\midrule
Temperature		                    & $^{\circ}$C        & 0.01  & Thermistor                           & Physical \\
Conductivity                        & $\mu$S/cm          & 0.01  & Four-Electrode Graphite Sensor       & Physical  \\
pH                                  & logarithmic (0-14) & 0.01  & Flowing-junction Reference Electrode & Physical \\
Turbidity                           & FNU                & 0.01  & Ion-Selective Electrode              & Physical \\
$\mathrm{Ca^{2+}}$                  & mg/l               & 0.1   & Ion-Selective Electrode              & Ions \\
$\mathrm{Cl^-}$                     & mg/l               & 0.1   & Ion-Selective Electrode              & Ions \\
$\mathrm{Na^+}$                     & mg/l               & 0.1   & Ion-Selective Electrode              & Ions \\
Blue-Green Algae (phycoerythrin)    & ppb                & 0.01  & Fluorometer                          & Biochemical \\
Blue-Green Algae (phycocyanin)      & ppb                & 0.01  & Fluorometer                          & BioChemical \\
CDOM                                & ppb                & 0.01  & Fluorometer                          & Biochemical \\
Chlorophyll A                       & ppb                & 0.01  & Fluorometer                          & BioChemical \\
Optical Brighteners                 & ppb                & 0.01  & Fluorometer                          & Chemical \\
Crude Oil                           & ppb                & 0.01  & Fluorometer                          & Chemical\\
\bottomrule
\end{tabularx}
% \noindent{\footnotesize{\textsuperscript{1} Tables may have a footer.}}
\end{table}

As shown in Table~\ref{tab:sensors}, the physical parameter and ion sensors are largely based on different configurations of electrodes while the chemical and biochemical parameters are all optically active substances enabling their determination by fluorometry. The pigments phycoerythrin and phycocyanin are used to determine blue-green algae content which together with chlorophyll-a enables us to asses the distribution of photosynthetic life in the pond. Additionally, we also measure the concentration of colored dissolved organic matter (CDOM) which impacts light penetration and serves as a primary source of bioavailable carbon. Crude oil (natural petroleum) and optical brightener concentrations are also measured with fluorometers and are relevant for identifying sources of industrial contamination or natural seepage.

@@

\begin{figure}[H]
\begin{adjustwidth}{-\extralength}{0cm}
\centering
\includegraphics[width=14cm]{paper/figures/materials-and-methods/pipeline-figure-2.pdf}
\end{adjustwidth}
\caption{Hyperspectral Image Processing: Hyperspectral data cubes are collected one scan-line at a time (left). By utilizing downwelling irradiance spectra we convert each pixel from spectral radiance to reflectance. By using orientation and position data from the on-board GPS and INS, we georeference each pixel to assign it a latitude and longitude on the ground. The final data product is a georectified hyperspectral reflectance data cube visualized at the right. \label{fig:hsi-pipeline}}
\end{figure}

The hyperspectral imager utilized in our robot team is in a so-called \textit{pushbroom} configuration, that is, each image captured by the drone is formed one scan-line at a time as the UAV flies. Each scan-line consists of 1600 pixels for which incoming light is diffracted into 462 wavelength bins. In the collection software, a regular cutoff of 1000 lines is chosen so that each resulting \textit{image} forms a array of size 462$\times$1600$\times$1000 called a \textit{hyperspectral data cube}. Initially, the captured spectra are in units of spectral radiance (measured in microflicks) however this does not account for the variability of incident light illuminating the water. To this end, we convert the hyperspectral data cubes into units of reflectance by utilizing the skyward-facing downwelling irradiance spectrometer (see Figure~\ref{fig:drone-components}). When the camera is normal to the water's surface, the reflectance is given by
\begin{equation}
    R(\lambda) = \pi L(\lambda)/E_d(\lambda)
\end{equation}
where $L$ is the spectral radiance, $E_d$ is the downwelling irradiance, and a factor of $\pi$ steradians results from assuming the water surface is Lambertian (diffuse) \cite{reflectance-conversion}. We will address the impact of this assumption in Section~\ref{sec:discussion}.

Having converted the hyperspectral data cube to units of reflectance, we must also georeference each pixel into a geographic coordinate system so that each image pixel can be assigned a latitude and longitude corresponding to the location on the ground from which it was sampled. During our three surveys, the UAV was flown at an altitude of ~50 m above the water. At this scale, the surface is essentially flat so that the hyperspectral data cube can be reliably georectified without the need for an on-board digital elevation map (DEM). We adopt the approach outlined in \cite{GeorectificationMuller, GeorectificationBaumker, GeorectificationMostafa} whereby each scan-line is georeferenced using the known field of view (30.8 $\deg$) together with the position and orientation of the UAV as provided by the on-board GPS INS. After a sequence of coordinate transformations, pixel coordinates are obtained in the relevant UTM zone (in meters). The resulting image is then re-sampled to a final output resolution. For these collections a resolution of 10 cm was utilized, however this can be adjusted to optimize the processing time and final resolution for real-time applications. Finally, the obtained pixel UTM coordinates are transformed to latitude and longitude for easy comparison with in-situ data collected by the USV. The final result is a georectified hyperspectral reflectance data cube. In Figure~\ref{fig:hsi-infographic} we visualize one such data cube, highlighting a selection of exemplar pixel spectra as well as the incident downwelling irradiance spectrum. A pseudo-color image is generated (plotted on the top of the data cube) to illustrate the scene.

@@

The above processing pipeline was implemented using the Julia programming language, a just-in-time compiled language with native multi-theading support \cite{julia-1}. By running this pipeline on the on board computer, we are able to process the collected hyperspectral data cubes in near real time. This feature is critical for time-sensitive applications where we need to quickly asses if an area is safe and can not afford to wait to download and post-process collected imagery \textit{after} a flight.

\subsection{Data Collection}

Data were collected using the robot team at a site in North Texas across three observation periods in the fall of 2020. For each collection period, the UAV first completed a wide survey of the pond capturing multiple hyperspectral data cubes. Subsequently, the the USV sampled across the pond, collecting in-situ reference measurements. Each of these reference measurements was then collocated with individual pixel spectra where the USV track overlapped with UAV pixels. To account for any time lag in the values measured by the in-situ instruments and to account for the USV size in comparison to the data cube's spatial resolution, each in-situ measurement is associated with a 3$\times$3 grid of HSI pixels, that is, a 30 cm $\times$ 30 cm square. These combined data form the tabular data set on which we train regression models; pixel spectra form input features and each separate USV sensor forms a distinct target variable.

Each collection was performed near solar noon to maximize the amount of incident sunlight illuminating the water, however, to account for any potential variation in lighting conditions between data cubes, we augment the training set with additional features including the drone viewing geometry (roll, pitch, heading) and solar illumination geometry (solar azimuth, solar elevation, solar zenith) as well as the total at-pixel intensity before reflectance conversion, the total downwelling intensity, and the drone altitude. Further feature engineering is performed to add additional spectral indices which utilize combinations of specific wavelength bands like the normalized difference vegetation index (NDVI), normalized difference water index (NDWI), simple ratio (SR), photochemical reflectance index (PRI), and more as outlined in \cite{envi_vegetation_indices, thenkabail2018hyperspectral,kaufman1992atmospherically, SpectralIndexWheat}. A comprehensive list of these added features is provided in supplementary table S1. The final dataset includes a total of 526 features (462 reflectance bands plus 64 additional features) with over 120,000 records.

@@

Using these tools, the training procedure for each model is as follows: First, the model is trained using 6-fold cross validation on the training set with default hyperparameter values. Feature importances for the trained model are then computed and the top 25 features are identified. A second, model is then trained using the same 6-fold cross validation scheme with \textit{only} these 25 most important features together with hyperparameter optimization using a random parameter search over the number of tees and sampling fraction. Model performance is evaluated by computing out-of-fold values for the coefficient of determination, as well as the
\begin{align}
    \text{RMSE} &= \sqrt{\frac{\sum\limits_{i=1}^N (\hat{y}_i-y_i)^2}{N}}, \\
    \text{MAE} &= \frac{\sum\limits_{i=1}^N \lvert \hat{y}_i - y_i \vert}{N},
\end{align}
where RMSE is the root-mean-square error, and MAE is the mean absolute error between true values $y_i$ and predictions $\hat{y}_i$.
